---
title: "INFO201 Final Project"
format: html
editor: visual
---

```{r}
# install.packages("dplyr")
library(dplyr)
library(ggplot2)
library(stringr)

```

# Gaming effect on mental health, anxious level, and working efficiency in US

## Importing Database 1: [The Effect of Screentime on the Mental Health of Children](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/1WWCA5)

This database analyzes 2020 NSCH data to examine the correlation between weekday screentime and mental health treatment among children aged 0–17. Findings show that increased screentime is associated with higher instances of mental health treatment, especially among adolescents (14–17 years old), where those needing treatment most often reported four or more hours of screentime.

Age: Integer, age Playing.Hours: Integer, hour participants spent on electronic device Productive.Time: Integer, hour participants spent on work or study Feel.After: String: "Good", "Tired", "Unmotivated" Mental.Harm: Catagory: No, Yes

```{r}
database_1 <- read.csv("First_database.csv")

# remove the line below later... just for visualizing the database
sample_n(database_1, 6) 
```

1.  filter the age (include only age over 18 years old)

```{r}
DB1_age_above_18 <- database_1 |> 
  filter(Age >= 18)

```

2.  ggplot bar chart to compare the playing hours for No mental harm and Yes mental harm

```{r}
DB1_age_above_18 |>
  ggplot(aes(x = Mental.Harm, y = Playing.Hours, fill = Mental.Harm)) +
  geom_bar(stat = "summary", fun = "mean") +   # using stat to calculate mean for playing hours
  labs(title = "Average Playing Hours and Mental Harm",
       x = "Mental Harm (Y/N)",
       y = "Average Playing Hours")
```

3.  ggplot bar chart to compare the playing hours for good and unmotivated

```{r}
DB1_age_above_18 |>
  ggplot(aes(x = Feel.After, y = Playing.Hours, fill = Feel.After)) +
  geom_bar(stat = "summary", fun = "mean") +   # using stat to calculate mean for playing hours
  labs(title = "Average Playing Hours and feeling after",
       x = "Feeling After Game",
       y = "Average Playing Hours")
```

4.  result conclusion 1:

## Importing Database 2: [Online Gaming Anxiety Data](https://www.kaggle.com/datasets/divyansh22/online-gaming-anxiety-data/data)

#### [Original study can be found here](https://osf.io/vnbxk/)

This dataset consists of data collected as a part of a survey among gamers worldwide. The questionnaire asked questions that psychologists generally ask people who are prone to anxiety, social phobia, and less to no life satisfaction. The questionnaire consists of several set of questions as asked as a part of psychological study. The original data was collated by Marian Sauter and Dejan Draschkow.

```{r}
database_2 <- read.csv("Second_database.csv")

# remove the line below later... just for visualizing the database
sample_n(database_2, 6) 
```

2.  Group by play style

```{r}
database_2 |>
  distinct(Playstyle)
```

# Playstyle VS GADT Histogram

```{r}

# Assigns a simple Multiplayer or Singleplayer label to each row
database_2 |>
  mutate(multi = case_when(
    (str_detect(tolower(Playstyle), "mix") | str_detect(tolower(Playstyle), "both")) ~ "Mix of Both",
    (str_detect(tolower(Playstyle), "friends") | str_detect(tolower(Playstyle), "One person")) ~ "Friends",
    TRUE ~ "Solo"
  )) |>
  
# Makes the graph
  ggplot(aes(x = GAD_T, fill = multi)) +
  geom_histogram() +
  facet_wrap(~multi)+
  labs(
    x = "Reported Anxiety score",
    y = "",
    fill = "Playstyle",
    title = "High anxiety levels seem to occur less in people who prefer to play alongside friends"
  )
```

3.  Summary statistics of anxious level questions

```{r}
multi_vs_single <- database_2 |>
  mutate(Playstyle_group = case_when(
    str_detect(tolower(Playstyle), "multiplayer") ~ "Multiplayer",
    TRUE ~ "Single/Others"
  )) |>
  group_by(Playstyle_group) |>
  summarise(
    count = n(),
    mean_GAD = mean(GAD_T, na.rm = TRUE),
    sd_GAD = sd(GAD_T, na.rm = TRUE),
    min_GAD = min(GAD_T, na.rm = TRUE),
    max_GAD = max(GAD_T, na.rm = TRUE)
  )

multi_vs_single
```

```{r}
data_for_test <- database_2 |>
  mutate(Playstyle_group = case_when(
    str_detect(tolower(Playstyle), "multiplayer") ~ "Multiplayer",
    TRUE ~ "Single/Others"
  )) |>
  select(Playstyle_group, GAD_T) |>
  filter(!is.na(GAD_T))

t_test_result <- t.test(GAD_T ~ Playstyle_group, data = data_for_test)
t_test_result
```
