---
title: "INFO201 Final Project"
format: html
editor: visual
---

```{r}
# install.packages("dplyr")
library(dplyr)
```

# Put our topic here!

## Importing Database 1: [The Effect of Screentime on the Mental Health of Children](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/1WWCA5)

This database analyzes 2020 NSCH data to examine the correlation between weekday screentime and mental health treatment among children aged 0–17. Findings show that increased screentime is associated with higher instances of mental health treatment, especially among adolescents (14–17 years old), where those needing treatment most often reported four or more hours of screentime.

```{r}
database_1 <- read.csv("First_database.csv")

# remove the line below later... just for visualizing the database
sample_n(database_1, 6) 
```

## Importing Database 2: [Online Gaming Anxiety Data](https://www.kaggle.com/datasets/divyansh22/online-gaming-anxiety-data/data)

#### [Original study can be found here](https://osf.io/vnbxk/)

This dataset consists of data collected as a part of a survey among gamers worldwide. The questionnaire asked questions that psychologists generally ask people who are prone to anxiety, social phobia, and less to no life satisfaction. The questionnaire consists of several set of questions as asked as a part of psychological study. The original data was collated by Marian Sauter and Dejan Draschkow.

```{r}
database_2 <- read.csv("Second_database.csv")

# remove the line below later... just for visualizing the database
sample_n(database_2, 6) 
```
